Quantifying the importance of a tree prior in Bayesian phylogenetics

### 2019-06-13 content.tex
Show result of example_6

\verb;pirouette; is written in the R programming language (\cite{R}).
The goal of \verb;pirouette; is to quantify the importance of a tree prior.
It does so by measuring the inference error made for a given reconstructed phylogeny, 
simulated under a (usually novel) diversification model.
We refer to the true prior that generated the given tree
as the 'generative tree prior' $\mathit{p_{G}}$.
Many tree priors have a parameter setting for which it reduces to a standard tree prior. For example, a protracted birth-death model reduces to a standard birth-death model when speciation-completion rate approaches infinity [i.e. rate $\lambda$ in \cite{etienne2014estimating}, eqs. (2b) and (2c)].
When benchmarking a novel tree prior, 
one will typically construct phylogenies 
for different combinations of the diversification model's parameters, 
to assess under which scenarios the inference error cannot be neglected. 
While we recommend many replicate simulations when assessing a novel tree prior, our examples contain only one replicate as they are for illustrative purposes only.

@@

\richel{New section about equal number of mutations}
From the twin tree, we simulate an alignment following the same
history as the twin tree, with the addition that 
the same total number of mutations have
taken place from root sequence to the sequences in the present.
This addition assures there is the same amount of historical information
stored in both the true and twin alignments.
We achieve this by simply simulating twin alignments until we
obtain one that has the desired number of mutations.

@@

We start from the tree generated by a non-standard (and unknown) tree prior, having six taxa and a crown age of ten:
\begin{lstlisting}[
  language = R, 
  floatplacement = ht,
  frame = single, 
  label = {lst:unknown_phylogeny},
  caption = A phylogeny generated by an unknown diversification model.
]
phylogeny  <- ape::read.tree(
  text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);"
)
\end{lstlisting}
\begin{figure}[H]
  \includegraphics[width=\textwidth]{example_2/true_tree.png}
  \caption{The tree derived from an unknown diversification process, 
    as created by listing~\ref{lst:unknown_phylogeny}.
  }
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=\textwidth]{example_3/errors.png}
  \caption{
    The inference error made 
    for both a generative tree prior and best candidate model
    compared with the error obtained for the twin tree.
    Here, the 'twin' tree shows the baseline inference error.
    Vertical dashed lines show the median error value per distribution.
  }
  \label{fig:example_3}
\end{figure}
\giovanni{
  I find the figure to be too confused. 
  What if we use instead different plots, 
  and in each of them we just show one inference model's results plus its twin counterpart?
}
\richel{Will do}

>>>>

\verb;pirouette; is written in the R programming language (\cite{R}).
The goal of \verb;pirouette; is to quantify the importance of a tree prior.
It does so by measuring the inference error made for a given reconstructed phylogeny, 
simulated under a (usually novel) diversification model.
We refer to the true prior that generated the given tree
as the 'generative tree prior' $\mathit{p_{G}}$.
Many tree priors have a parameter setting for which it reduces to a standard tree prior. For example, a protracted birth-death model reduces to a standard birth-death model when speciation-completion rate approaches infinity [i.e. rate $\lambda$ in \cite{etienne2014estimating}, eqs.~(2b) and (2c)].
When benchmarking a novel tree prior, 
one will typically construct phylogenies 
for different combinations of the diversification model's parameters, 
to assess under which scenarios the inference error cannot be neglected. 
While we recommend many replicate simulations when assessing a novel tree prior, our examples contain only one replicate as they are for illustrative purposes only.

@@

From the twin tree, we simulate an alignment following the same
history as the twin tree, with the addition that 
the same total number of mutations have
taken place from root sequence to the sequences in the present.
This addition assures there is the same amount of historical information
stored in both the true and twin alignments.
We achieve this by simply simulating twin alignments until we
obtain one that has the desired number of mutations.

@@

We start from the tree generated by a non-standard (and unknown) tree prior, having six taxa and a crown age of ten:
\begin{lstlisting}[
  language = R, 
  floatplacement = ht,
  frame = single, 
  label = {lst:unknown_phylogeny},
  caption = A phylogeny generated by an unknown diversification model.
]
phylogeny  <- ape::read.tree(
  text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);"
)
\end{lstlisting}
\begin{figure}[H]
  \includegraphics[width=\textwidth]{example_6/true_tree.png}
  \caption{The tree derived from an unknown diversification process, 
    as created by listing~\ref{lst:unknown_phylogeny}.
  }
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=\textwidth]{example_6/errors.png}
  \caption{
    The inference error made 
    for both a generative tree prior and best candidate model
    compared with the error obtained for the twin tree.
    Here, the 'twin' tree shows the baseline inference error.
    Vertical dashed lines show the median error value per distribution.
  }
  \label{fig:example_3}
\end{figure}
\giovanni{
  I find the figure to be too confused. 
  What if we use instead different plots, 
  and in each of them we just show one inference model's results plus its twin counterpart?
}
\richel{Will do}
